---
title: "Homework3-2"
format: html
editor: visual
---

## Task 1

### 1) If your working directory is myfolder/homework/, what relative path would you specify to get the file located at myfolder/MyData.csv?

read.csv(myfolder/Mydata.csv)

### 2) What are the major benefits of using R projects?

The major benefit of using R projects is the usage of relative paths.

### 3) What is git and what is github?

Git is open source and installed locally, whereas github is on a website database.

### 4) What are the two main differences between a tibble and a data.frame?

Tibbles have all variable types, and cannot partially match.

### 5) Rewrite the following nested function call using baseR’s chaining operator:

```{r}
library(dplyr)
iris_tbl <- as_tibble(iris) |>
  select(starts_with("Petal"), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```

### 6) What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?

Long format data is when one row represents one observation, whereas wide is one column represents one observation. For statistical analysis, long format data is preferred.

## Task 2 - Glass Data

```{r}
library (tidyverse)
library(readr)
library(dplyr)
library(Lahman)
#1

#read in the raw data file named glass data, and put in column names.
glass_data <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/glass.data",
                          delim = ",",
                          col_names = c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe"))

glass_data

#2

#overwrite last column with character strings 
glass_data_tbl <- as_tibble(glass_data)
glass_data_tbl |>
  mutate(desc = ifelse(glass_data_tbl$X11 == 1, "building_windows_float_processed",
              ifelse(glass_data_tbl$X11 == 2, "bulding_windows_non_float_processed",
              ifelse(glass_data_tbl$X11 == 3, "vehicle_windows_float_processed",
              ifelse(glass_data_tbl$X11 == 4, "vehicle_windows_non_float_processed",
              ifelse(glass_data_tbl$X11 == 5, "containers",
              ifelse(glass_data_tbl$X11 == 6, "tableware", "headlamps"))))))) |>
#3
  #continue the chain and filter based on requirements.
  filter(Fe < 0.2, desc %in% c("tableware", "headlamps"))

print(glass_data_tbl)

```

## Task 2 - Yeast Data

```{r}
library(readr)
#1 - Read in raw delimited data and create column names based on HW requirements.
columnnames = c("seq_name", 
                "mcg",
                "gvh", 
                "alm", 
                "mit", 
                "erl",
                "pox",
                "vac",
                "nuc",
                "class")
yeast_data <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/yeast.data",
                         delim = "  ",
                         col_names = columnnames)
print(yeast_data)
yeast_data_tbl <- as_tibble(yeast_data)
print(yeast_data_tbl)

#2 - Remove seq_name and nuc columns from tibble.
yeast_data_tbl |>
  select(-seq_name, -nuc) |>
#add mean and median columns corresponding to each numeric variable.
  mutate(across(where(is.numeric), list(mean = mean, median = median), na.rm = TRUE, .names = "{.col}_{.fn}"))
  
  
```

#Task 2: Combining Excel and Delimited Data

```{r}

#1
library(readxl)
#import raw excel data, just the first sheet.
white_wine <- read_excel("white-wine.xlsx", 
                         sheet = excel_sheets("white-wine.xlsx")[1])
white_wine_tbl <- as_tibble(white_wine)
view(white_wine_tbl)

print(white_wine_tbl)


#2
#import second sheet of raw excel data with all variable names and info.
white_wine_2nd <- read_excel("white-wine.xlsx",
                             sheet = excel_sheets("white-wine.xlsx")[2])
#make column names of RData object as variables from 2nd sheet.
colnames(white_wine) <- white_wine_2nd$Variables

#3
#add additional column as per HW requirements.
white_wine_tbl$Type = c("White")

white_wine$Type = c("White")



#4 - do same as above to red wine raw delimited file.
red_wine <- read_delim("https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
                       delim = ";")
colnames(red_wine) <- white_wine_2nd$Variables  

red_wine$Type = c("Red")

#5 - combine the two datasets together
wines_overall <- bind_rows(white_wine, red_wine)
print(wines_overall)

#6 through 9
library(dplyr)
filtered_wines_overall <- wines_overall |>
  #arrange quality variable from descending order, largest to smallest.
  arrange(desc(wines_overall$quality)) |>
  #filter data with quality < 6.5 and alcohol < 132.
  filter(wines_overall$quality > 6.5, wines_overall$alcohol < 132) |>
  #Select only certain variables.
  select(contains("acid"),contains("alcohol"), "Type", "quality") |>
  group_by(quality) |>
  #add mean and standard deviation of alcohol variable for each quality setting.  To do this, need to first use group_by function as shown in one line above.
  mutate(wine_alcohol_mean = mean(alcohol),
         wine_alcohol_sd = sd(alcohol))

  print(filtered_wines_overall)
```

#Task 3 - Database Practice

```{r}
library(DBI)
library(RSQLite)

#1 - Connect to database
lahman_db <- dbConnect(RSQLite::SQLite(), "lahman.db")
dbListTables(lahman_db)

#2 -  return all data from Teams table in the year 2015.
tbl(lahman_db, "Teams") |>
  filter(yearID == 2015)

#3 -  Use SQL to do the same as shown in question 2.
lahman_db2 <- dbConnect(RSQLite::SQLite(), "lahman.db")

tbl(lahman_db2, sql(
  "SELECT 'yearID', 'playerID', 'teamID', 'lgID' 
  FROM 'Teams' 
  WHERE ('yearID' = 2015.0)")
  )
  
#4 - Return all players from Hall of Fame table.
lahman_db_modified <- tbl(lahman_db, "HallOfFame") |>
  select(playerID, yearID, category) |>
print(lahman_db_modified)

#5 - Combine People table with previously modified HallOfFame table.
lahman_db_join <- tbl(lahman_db, "People") |>
left_join(lahman_db_modified, tbl(lahman_db,"People"),
          by = join_by(playerID == playerID)) |>
  collect() |>
  select(playerID, yearID, nameFirst, nameLast)
print(lahman_db_join)

#6 - Using chaining to select certain variables in table, group by one variable, and summarize win/loss ratio.
lahman_db_managers <- tbl(lahman_db, "Managers") |>
      select(playerID, G, W, L) |>
      group_by(playerID) |>
      summarize(G_managed = sum(G, na.rm = TRUE),
             Total_W = sum(W, na.rm = TRUE),
             Total_L = sum(L, na.rm = TRUE)) |>
      collect() |>
      mutate(win_loss_percentage = Total_W/G_managed) |>
      arrange(desc(win_loss_percentage))
print(lahman_db_managers)

#7 - returning all variables from last two questions.
lahman_total_join <- full_join(lahman_db_managers, lahman_db_join, 
                               by = join_by(playerID == playerID)) |>
                      select(everything())
                                 
print(lahman_total_join)                               

```
